Supporting Information: The Regional Effects of Marine Protected Areas

### 2020-04-21 documents/reviewer-responses.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reviewer Responses

We thank the two reviewers and the Editor for their constructive responses. We have fundamentally restructured the paper to place much greater emphasis on the empirical results from the Channel Islands MPA network. This is to our knowledge the first empirical assessment of the regional effects of a large and iconic MPA network on a wide assemblage of commercially and economically important species. Our inability to estimate a clear effect in such a well-enforced and well-studied local is we feel a vitally important "null-result" both for scientists and MPA managers to be aware of. As such we feel that it warrants publication in PNAS.

While our results are more focused on the empirical results, we explicitly link our results to the large body of MPA theoretical work, using it to help illustrate why we might expect such a null result and help managers understand when they might be able to detect an effect. As MPAs are used as a core tool of marine resource management, our results provide a useful update to our empirical expectations for MPAs, that to date have been based almost entirely on inside-outside MPA response ratios. We hope you find the updated version of our manuscript improved thanks to the inclusions of the careful points raised by the prior revision of this paper.

### 2020-04-21 documents/ovando-regional-effects-of-mpas-si-v2.Rmd
dual run on squid

---
title: "Supporting Information: The Regional Effects of Marine Protected Areas"
author:  Daniel Ovando,Jennifer E. Caselle,Christopher Costello,Olivier Deschenes,Steven D. Gaines,Ray Hilborn,Owen Liu
output: 
  bookdown::pdf_document2:
    keep_tex: true
  pdf_document: default
bibliography: My Library.bib
header-includes:
    - \usepackage{longtable}
    - \usepackage{booktabs}
    - \usepackage{eso-pic}
    - \usepackage{graphicx}
    - \usepackage[left]{lineno}
    - \usepackage{xcolor}
    - \linenumbers
params:
  run_name: ["v5.0"]
  sim_years: [50]
  burn_years: [20]
  num_patches: [25]
always_allow_html: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE)

library(sf)
library(ggmap)
library(viridis)
library(hrbrthemes)
library(scales)
library(patchwork)
library(rpart)
library(extrafont)
library(caret)
library(ggsci)
library(rEDM)
library(tidyverse)
library(Synth)
library(spasm)
extrafont::loadfonts()
functions <- list.files(here::here("functions"))

@@

# sim_years <- 50
# 
# burn_years <- 20
# 
# num_patches <- 25
# 
# run_name <- "v5.0"

@@

load(file = here::here("results",run_name, "model_runs.Rdata"))

load(file = here::here("results",run_name,"processed_grid.Rdata"))

load(file.path(run_dir,"abundance_data.Rdata"))

@@

@@

facet_labels <- c(
  mpa_size = "Range in MPA",
  depletion = "Without-MPA Depletion"
)

channel_islands <- readRDS(here::here("data","channel_islands_map.RDS"))

ca_mpas <- sf::st_read(here::here("data","MPA_CA_Existing_160301")) %>%
  rmapshaper::ms_simplify() %>%
  sf::st_transform(crs = 4326)

zissou_theme <-
  theme_ipsum(
    base_size = 12,
    axis_title_size = 12,
    strip_text_size = 12,
    base_family = "Helvetica"
  )

zissou_theme <- zissou_theme +
  theme(   panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank())

@@

theme_set(zissou_theme)

models_worked <- model_runs$tmb_fit %>% map("error") %>% map_lgl(is_null)

model_runs <- model_runs %>%
  filter(models_worked) %>%
  mutate(tmb_fit = map(tmb_fit,"result")) %>%
  mutate(processed_fits = map(tmb_fit, process_fits)) %>%
  mutate(did_plot = map(processed_fits, "did_plot"))

@@

fitted_data <- base_run$data[[1]]

zissou_fit <- base_run$tmb_fit[[1]]

report <- base_run$tmb_fit[[1]]

@@

zissou_fit$zissou_estimates <- zissou_fit$zissou_estimates %>%
  mutate(variable = str_replace_all(variable, "(\\.).*$",""))

report$zissou_estimates <- report$zissou_estimates %>% 
    mutate(variable = str_replace_all(variable, "(\\.).*$",""))

seen_non_nested_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seen_non_nested_betas")) %>%
  rename(group = variable) %>%
  mutate(variable  = zissou_fit$zissou_data$x_seen_non_nested %>% colnames())

seeing_non_nested_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seeing_non_nested_betas")) %>%
  rename(group = variable) %>%
  mutate(variable  = zissou_fit$zissou_data$x_seen_non_nested %>% colnames())

seen_year_species_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seen_year_species_betas")) %>%
  rename(group = variable) %>%
  mutate(variable = zissou_fit$zissou_data$x_seen_year_species %>% colnames())

seeing_year_species_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seeing_year_species_betas")) %>%
  # filter(variable == "seeing_year_species_betas") %>%
  rename(group = variable) %>%
  mutate(variable = zissou_fit$zissou_data$x_seen_year_species %>% colnames())

seen_region_cluster_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seen_region_cluster_betas")) %>%
  # filter(variable == "seen_region_cluster_betas") %>%
  rename(group = variable) %>%
  mutate(variable = zissou_fit$zissou_data$x_seen_region_cluster %>% colnames())

seeing_region_cluster_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seeing_region_cluster_betas")) %>%
  # filter(variable == "seeing_region_cluster_betas") %>%
  rename(group = variable) %>%
  mutate(variable = zissou_fit$zissou_data$x_seen_region_cluster %>% colnames())

did_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "mpa_effect")) %>%
  # filter(variable == "mpa_effect") %>%
  mutate(group = variable) %>%
  mutate(year = zissou_fit$did_data$year %>% unique())

 betas <- bind_rows(
   seen_non_nested_betas,
   seeing_non_nested_betas,
   seen_year_species_betas,
   seeing_year_species_betas,
   seen_region_cluster_betas,
   seeing_region_cluster_betas,
   did_betas %>% select(-year)
 ) %>%
   as_data_frame()

 non_nested_beta_plot <- betas %>%
   filter(str_detect(group, "non_nested")) %>%
   ggplot() +
   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
   geom_pointrange(aes(x = variable,
                       y = estimate,
                       ymin = lower,
                       ymax = upper)) +
   facet_wrap(~group) +
   coord_flip() +
   theme(axis.text.y = element_text(size = 10))

 year_species_effects_plots <- betas %>%
   filter(str_detect(group, "year_species_betas")) %>%
   mutate(year = str_replace_all(variable,"\\D","") %>% as.numeric()) %>%
   mutate(classcode = str_split(variable,'-', simplify = T)[,2]) %>%
   ggplot() +
   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
   geom_ribbon(aes(x = year, ymin = lower, ymax = upper, fill = classcode), alpha = 0.25) +
   geom_line(aes(x = year, y = estimate, color = classcode)) +
   facet_wrap(~group)

  region_cluster_plots <- betas %>%
   filter(str_detect(group, "region_cluster_betas")) %>%
   mutate(cluster = str_replace_all(variable,"\\D","")) %>%
   mutate(region = str_split(variable,'-', simplify = T)[,3]) %>%
   ggplot() +
    geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
    geom_pointrange(aes(x = region,
                        y = estimate,
                        ymin = lower,
                        ymax = upper, color = cluster)) +
    facet_wrap(~group)

  sim_plot <- outcomes %>%
    mutate(rsize = plyr::round_any(mpa_size, 0.25),
           reff = plyr::round_any(mpa_effect,0.1)) %>%
    filter(rsize == 0.25) %>%
    mutate(year = years_protected + 2001) %>%
    filter(year %in% did_betas$year) %>% 
    group_by(year, reff) %>% 
    count() %>% 
    group_by(year) %>% 
    mutate(sc = n / sum(n))

  did_plot <- did_betas %>%
    ggplot() +
    geom_tile(data = sim_plot, aes(year, reff, fill = sc), alpha = 0.75) +
    geom_vline(aes(xintercept = 2003), color = 'red', linetype = 2, size = 2) +
    geom_hline(aes(yintercept = 0)) +
    geom_pointrange(aes(
      year,
      y = exp(estimate) - 1,
      ymin = exp(lower) - 1,
      ymax = exp(upper) - 1
    ),
    size = 1.5
    ) +
    labs(x = "Year", y = "Targeted Trend Divergence") +
    scale_fill_viridis(guide = hgc,
                       name = "% of Annual Sims",
                       labels = scales::percent_format(accuracy = 1),
                       option = "plasma") + 
    theme(legend.position = "top")

top_species <- base_run$data[[1]]$classcode %>% unique()

cip_data <- pisco_data %>%
  left_join(site_data, by = c("site","side")) %>%
  filter(is.na(eventual_mpa) == F) %>%
  filter(region %in% c("ANA", "SCI","SRI",'SMI'),
         classcode %in% top_species)

# process outcomes

filter_summary <- fitted_data %>% 
  select(classcode, targeted) %>% 
  unique() %>% 
  group_by(targeted) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(targeted = ifelse(targeted == 1,"yes","no")) %>% 
  spread(targeted, n)

@@

```{r pkg-info, results = "asis"}
table <- devtools::package_info() %>% 
  select(package, loadedversion, date,source)

@@

>>>>

---
title: "Supporting Information: The Regional Effects of Marine Protected Areas"
author:  Daniel Ovando,Jennifer E. Caselle,Christopher Costello,Olivier Deschenes,Steven D. Gaines,Ray Hilborn,Owen Liu
output: 
  bookdown::pdf_document2:
    keep_tex: true
  pdf_document: default
bibliography: My Library.bib
header-includes:
    - \usepackage{longtable}
    - \usepackage{booktabs}
    - \usepackage{eso-pic}
    - \usepackage{graphicx}
    - \usepackage[left]{lineno}
    - \usepackage{xcolor}
    - \linenumbers
params:
  run_name: ["v6.0"]
  sim_years: [50]
  burn_years: [20]
  num_patches: [25]
always_allow_html: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
                      cache = FALSE, fig.width = 3.42, fig.asp = 1.5)

```{r, include=FALSE}
library(sf)
library(ggmap)
library(viridis)
library(hrbrthemes)
library(scales)
library(patchwork)
library(rpart)
library(extrafont)
library(caret)
library(ggsci)
library(rEDM)
library(tidyverse)
library(ggtext)
library(rstan)
library(rstanarm)
library(bayesplot)
extrafont::loadfonts()
rstan_options(auto_write = TRUE)

functions <- list.files(here::here("functions"))

@@

# sim_years <- 50
# 
# burn_years <- 20
# 
# num_patches <- 25

# run_name <- "v5.0"

width_1col <- 3.42 # inches

asp_1col <- 1.25

width_2col <- 7 # inches

asp_2col <-  .75

dpi <- 900

theme_1col <- hrbrthemes::theme_ipsum(base_size = 11,
                                      axis_title_size = 12)

theme_2col <- hrbrthemes::theme_ipsum(base_size = 12,
                                        axis_title_size = 14)

@@

load(file.path(run_dir, 'abundance_data.Rdata'))

model_runs <- read_rds( file.path(run_dir,"did_fits.rds"))

# load(file = here::here("results",run_name, "model_runs.Rdata"))

processed_grid <-  read_rds(file.path(run_dir,"filtered_processed_grid.rds"))

load(file = file.path(run_dir ,"sim_grid.Rdata"))

load(file = file.path(run_dir,"plots.RData"))

valgrid <- read_rds(file.path(run_dir, "valgrid.rds"))

cdfw_catches <-
  read_csv(file = here::here("data", 'cdfw-catches.csv')) %>%
  group_by(sci_name, year) %>%
  summarise(catch = sum(pounds_caught, na.rm = T)) %>%
  left_join(
    life_history_data %>% select(taxa, classcode, commonname) %>% mutate(taxa = tolower(taxa)),
    by = c("sci_name" = "taxa")
  ) %>%
  filter(!is.na(classcode))

# cdfw_catches %>%
#   group_by(commonname, classcode, sci_name) %>%
#   summarise(catch =sum(catch, na.rm = TRUE)) %>%
#   mutate(in_analysis = classcode %in% unique(cip_data$classcode)) %>%
#   View()

@@

outcomes <- outcomes %>% 
  filter(!is.na(mpa_effect))

density_ratios <- density_ratios %>% 
  filter(!is.na(biased_density_ratio))

@@

facet_labels <- c(
  mpa_size = "Range in MPA",
  depletion = "Depletion"
)

@@

theme_set(theme_1col)

# models_worked <- model_runs$tmb_fit %>% map("error") %>% map_lgl(is_null)

# model_runs <- model_runs %>%
#   filter(models_worked) %>%
#   mutate(tmb_fit = map(tmb_fit,"result")) %>%
#   mutate(processed_fits = map(tmb_fit, process_fits)) %>%
#   mutate(did_plot = map(processed_fits, "did_plot"))

@@

used_data <- abundance_data$data[abundance_data$data_source == "pisco"][[1]]

did_data <- model_runs$did_fit[model_runs$data_source == "pisco" &
                                 model_runs$var_names == "pisco_a" &
                                 model_runs$data_to_use == "all"][[1]]$did_data

did_fit <- model_runs$did_fit[model_runs$data_source == "pisco" &
                                 model_runs$var_names == "pisco_a" &
                                 model_runs$data_to_use == "all"][[1]]$did_reg

did_results <- model_runs$did_fit[model_runs$data_source == "pisco" &
                                 model_runs$var_names == "pisco_a" &
                                 model_runs$data_to_use == "all"][[1]]$did_results

@@

top_species <- used_data$classcode %>% unique()

life_summaries <- life_history_data %>% 
  select(classcode, tm, vbgf.k, vbgf.linf) %>% 
  rename(age_mature = tm, vbk = vbgf.k, linf = vbgf.linf) %>% 
  unique() %>% 
  filter(classcode %in% top_species) %>% 
  select(-classcode) %>% 
  gather(variable,value) %>% 
  group_by(variable) %>% 
  summarise(min_val = min(value, na.rm = TRUE),
         max_val = max(value, na.rm = TRUE))

 sim_life_history <- sim_grid %>%
    mutate(
      age_mature = map_dbl(fish, "age_mature"),
      vbk = map_dbl(fish, "vbk"),
      linf = map_dbl(fish, "linf")
    ) %>%
    select(scientific_name, age_mature, vbk, linf) %>%
    unique() %>% 
   filter(age_mature > life_summaries$min_val[life_summaries$variable == "age_mature"],
          age_mature < life_summaries$max_val[life_summaries$variable == "age_mature"],
          vbk > life_summaries$min_val[life_summaries$variable == "vbk"],
          vbk < life_summaries$max_val[life_summaries$variable == "vbk"],
          linf > life_summaries$min_val[life_summaries$variable == "linf"],
          linf < life_summaries$max_val[life_summaries$variable == "linf"])

 year_bins <-
  seq(2003, max(pisco_data$year) + 1, by = 3)

 sim_mpa_effects <- outcomes %>% 
  filter(set == "ci",
         final_depletion > 0,
         f_v_m > .5) %>% 
  mutate(year = years_protected + 2002) %>% 
  filter(year <= max(pisco_data$year)) %>% 
  mutate(binned_year = cut(year, year_bins, include.lowest = FALSE))  %>% 
  filter(year >= 2003,
         !is.na(binned_year)) %>% 
  mutate(mpa_effect = pmin(mpa_effect,3))
```

@@

```{r pkg-info, results = "asis", eval=FALSE}
table <- devtools::package_info() %>% 
  select(package, loadedversion, date,source)

@@

## PISCO Data

All fish data used in the primary difference-in-difference model were collected by PISCO. The dive transect survey methods are described in @caselle2015, provided below for ease of reference

"Fish assemblages were surveyed annually as part of a long-term monitoring program conducted by the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) using standard underwater visual belt survey methods (www.piscoweb.org9). We analyzed data from 47 PISCO sites at the northern Channel Islands that were sampled annually from at least 2003 to 2012. We also excluded 3 sites on Anacapa where a much older MPA had already been established in 1978. MPAs on each island were sampled annually during June-October and we surveyed multiple sites inside and outside of any individual MPA. Details of MPA characteristics such as size and coastline extent are given in Hamilton et al.9. At each site, we conducted 8 to 12 fish transects that measured 30×2×2m at multiple levels in the water column: benthic, midwater, and kelp canopy (when present). Transects are laid out in a stratified random design, with multiple nonpermanent transects located in fixed strata (i.e., outer, middle, and inner edges of the reef). At each level in the water column, one SCUBA diver per transect counted and estimated the sizes of all fish to the nearest centimeter (total length), excluding small cryptic fishes"
- @caselle2015

We take a number of steps to translate the raw transect data into the total biomass densities used in this study. For the default run, we only include species that were observed at least once for at least 15 of the 18 years of available data. We also exclude "young of the year" observations due the challenges of correctly identifying and measuring these individuals. We omitted data from 1999 due to changes in the sampling procedures that occurred after 1999. Per recommendations from PISCO staff we omit observations from the canopy level of the transects (leaving the middle, bottom, and middle canopy levels).

Once zeros have been introduced to the database, we convert positive observations of fish from numbers to biomass. Each observation in the raw database lists the species, the number of individuals seen, and the size of those individuals (either as one value or as a mininum and maximum size for the group seen). PISCO staff compiled allometric information used to convert lengths to expected weights. For each observation then, we convert the observed lengths to weights per these relationships (accounting for variations in length types such as standard vs. total length). When minimum and maximum ranges were reported, we drew a number of samples equal to the number of observed fish in that group from a uniform distribution spanning the minimum and maximum reported size in that group. We assume all length-to-weight conversions are constant and deterministic.

For each species at each transect, we then calculate the biomass density of that transect as the sum of the observed biomass divided by the transect area. We then average the biomass densities for each species across all the transects at a given site, and lastly sum these mean species biomass densities to achieve the total mean species biomass at the site level.

We include several additional sources of data in our regression analysis. Temperature readings are included from the PISCO data for each transect. We also include PISCO data on the estimated surge and visibility. We augmented these data with information on kelp cover over time from the Santa Barbara Channel Long Term Ecological Research Network [@lter2017]. We used a k-nearest neighbors algorithm to fill in missing kelp observations, and matched the interpolated kelp data to the PISCO data at the resolution of year-month-site (Fig.\@ref(fig:kelp-plot)). We include a variable capturing the mean cumulative number of observations across all observers conducting transects, in an effort to control for evoloving observer skill.

<!-- Temperature data were augmented with data from `FishLife` [@thorson2017c] to include the estimated preferred temperature for a given species, so that we can include deviations from the preferred temperature envelope as a predictor in the model. This allows different temperatures to have different effects on each species (and is less computationally intensive than estimating species-temperature slopes) (Fig.\@ref(fig:tempdev-plot)).  -->

used_data %>% 
  filter(targeted == 1) %>% 
  group_by(commonname, year) %>% 
  summarise(catches = unique(lag_catch)) %>% 
  group_by(commonname) %>% 
  mutate(total_catches = sum(catches)) %>% 
  ungroup() %>% 
  filter(total_catches > 0) %>% 
  ggplot(aes(year, catches)) + 
  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
  geom_line() + 
  facet_wrap(~commonname, scales = "free_y") + 
  theme_minimal() + 
  labs(x = "Year", y = "Catch (lbs)")

used_data %>% 
  group_by(region,year) %>% 
  summarise(mean_kelp = mean(interp_kelp, na.rm = TRUE)) %>% 
  ggplot(aes(year, (mean_kelp), color = region)) + 
  geom_line() + 
  labs(x = "Year", y = "Mean Kelp Biomass")

species_data <- used_data %>% 
  select(classcode, commonname, taxa, targeted) %>% 
  unique() %>% 
  arrange(desc(targeted)) %>% 
  mutate(targeted = targeted == 1) %>% 
  rename(`Common Name` = commonname, 
         `Scientific Name` = taxa,
         `Targeted?` = targeted)

## Difference-in-Difference Model

The difference-in-difference (DiD) regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species minus the same difference for non-targeted species in the Channel Islands.

The simplified form of this model is

\begin{equation}
  d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + \mathbf{B^cX_i} + \mathbf{B^sS_i}},shape, scale)
\label{eq:did}
\end{equation}

To provide greater detail on this process, we first conduct the data pre-processing described earlier. From there, we aggregated data to the level of total biomass density of targeted and non-targeted species at each site each year. As such, our base model estiamtes the effect of the MPAs on total biomass density of targeted species, though we also explore the effect on alternative specifications such as mean biomass of targeted species. the model was then fit using the `rstanarm` package in R (5000 iterations, 2500 warmup, 4 chains, adapt_delta = 0.85).

The intercept prior was was determined using the `rstanarm` `autoscale` function. For the non-intercept terms, we manually set a normal(0,2) prior. This implies that coefficeints such as the MPA effect have a prior that provides support for an effect size centered on zero with a range of roughly -4 to 4. Since the covariates are centered and scaled, and the dependent variable is on the log-scale, this is an extremely diffuse prior.

The full table of covariates and their posterior means and 89% credible interval can be found in Table.S\@ref(tab:did-coefs).

```{r did-coefs, results = "asis"}

did_fit %>% 
  broom::tidy(intervals = TRUE, prob = 0.89) %>% 
  select(term, estimate, lower, upper) %>% 
  rename(mean = estimate, lower_89th_percentile = lower, upper_89th_percentile= upper) %>% 
  knitr::kable(
      digits = 2,
            caption = "Posterior means and 89% credible interval for key model coefficients",
             longtable = TRUE,
             booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
```

```{r prior-plot, fig.cap="Prior and posterior distributions of the estimated MPA effect over time"}

priors <- tibble(year = unique(did_results$year)) %>% 
  mutate(prior = list(rnorm(500,0,2))) %>% 
  unnest(cols = prior)

did_results %>% 
  filter(between(prank, 0.05, 0.95)) %>% 
  ggplot() + 
  geom_density(aes(did, fill = "Posterior", y = ..scaled..), color = "transparent", alpha = 0.75) + 
  geom_density(data = priors, aes(prior, fill = "prior", y = ..scaled..), color = "transparent",
               alpha = 0.5) +
  facet_wrap(~year) + 
  scale_fill_discrete(name = '') + 
  scale_y_continuous(name = "Scaled Density")

We include a series of standard visual diagnostic plot for Bayesian models below, with brief descriptions of the key outcome of the diagnostic included in the figure captions. The model had 0 divergences or max tree depth saturations.

```{r, echo = FALSE}

rstan::check_hmc_diagnostics(did_fit$stanfit)

```{r rhat-plot, fig.cap="Histogram of potential scale reduction statistic Rhat. All values are below 1.05, indicating there is not evidence of chain convergence failure"}

data.frame(rhat = bayesplot::rhat(did_fit)) %>% 
             ggplot(aes(rhat)) + 
             geom_histogram() + 
             geom_vline(aes(xintercept = 1.05))

```{r energy-plot, fig.cap = "HMC energy diagnostic plots. Closely matching histograms shows there is no evidence of excessively fat tails"}

mcmc_nuts_energy(nuts_params(did_fit))

```{r}
y <- (did_fit$y)

yrep <- (rstanarm::posterior_predict(did_fit, draws = 500))
```

```{r, fig.cap = "Posterior Predictive distribution of total biomass density (blue lines) and observed distribution of total biomass density (red line)"}

ppc_dens_overlay(y = y,
                 yrep = yrep) +
  scale_x_continuous(name = "Log(Total Biomass Density)") +
  scale_colour_discrete(labels = c("Observed", "Posterior Predictive"),
                       name = '')

```{r mean_test_plot, fig.cap = "Posterior predictive mean (distribution) and empirical mean of total biomass density. Posterior predictive mean adequately recovers empirical mean."}

ppc_stat(y,  yrep, stat = "mean") 
```

```{r mean_test_plot, fig.cap = "Posterior predictive sum (distribution) and empirical sum of total biomass density. Posterior predictive sum adequately recovers empirical sum."}

ppc_stat(y,  yrep, stat = "sum") 
```

```{r sd_test_plot, fig.cap = "Posterior predictive standard deviation (distribution) and empirical standard deviation of total biomass density. Posterior predictive standard deviation adequately recovers empirical standard deviation"}
ppc_stat(y,  yrep, stat = "sd")

```{r min_test_plot, fig.cap = "Posterior predictive minimum (distribution) and empirical minimum of log total biomass density. Model slightly overpredicts minimum value in the data"}

ppc_stat(y,  yrep, stat = "min") 
```

```{r max_test_plot, fig.cap = "Posterior predictive maximum (distribution) and empirical maximum of total biomass density. Model underestimates true maximum of the data."}
ppc_stat(y,  yrep, stat = "max")

```{r resid-plot,fig.cap = "Posterior predictive total biomass densities plotted against mean posterior predictive residuals" }
ppc_error_scatter_avg_vs_x(y, yrep, colMeans(yrep)) + 
  scale_x_continuous(name = "Predicted total biomass density") + 
  scale_y_continuous(name = "Mean Posterior Predictive Residuals") + 
  geom_hline(yintercept = 0)
```

```{r resid-hist-plot, fig.cap = "Histograms of residuals for targeted and non-targeted obesrvations."}

temp <- did_data %>% 
  mutate(targ = ifelse(targeted == 1, "Targeted","Non-Targeted"))
ppc_error_hist_grouped(y, yrep = yrep[1:2,], group = temp$targ) + 
  geom_vline(aes(xintercept = 0)) + 
  scale_x_continuous(name = "Residuals")
```

### Additional Difference-in-difference Runs

We include a variety of additional model runs designed to explore the sensitivity of our key results to various assumptions and data processing steps included in our base results. Description and key message are provided inside figure captions.

Our base run uses data collected by PISCO. As a robustness check to our main results, we repeated our analysis utilizing data provided by the [Kelp Forest Monitoring Program (KFM)](https://science.nature.nps.gov/im/units/medn/monitor/kelpforest.cfm) conducted in the Channel Islands.

```{r, fig.cap="Model fit to PISCO data but only using variables also available for KFM data", results="hide"}

val_runs <- model_runs %>% 
  mutate(did_results = map(did_fit,"did_results"))

ggplot() +
 tidybayes::stat_halfeye(
    data = val_runs$did_results[[2]],
    aes(year, did, color = "Empirical Estimate"),
    .width = c(0.5, 0.95),
    slab_alpha = 0.95  ) +
  scale_y_continuous(name = "MPA Effect") +
  scale_x_discrete(name = element_blank()) +
  theme(legend.position = "top") +
  scale_fill_brewer(name = "Simulation Quantiles") + 
  scale_color_manual(values = "red", name = '') + 
  geom_hline(yintercept = 0)

```{r, fig.cap="Model fit to KFM data. Results are more uncertain than using PISCO data, but follow the same trend estimated from the PISCO data."}

ggplot() +
 tidybayes::stat_halfeye(
    data = val_runs$did_results[[3]],
    aes(year, did, color = "Empirical Estimate"),
    .width = c(0.5, 0.95),
    slab_alpha = 0.95  ) +
  scale_y_continuous(name = "MPA Effect") +
  scale_x_discrete(name = element_blank()) +
  theme(legend.position = "top") +
  scale_fill_brewer(name = "Simulation Quantiles") + 
  scale_color_manual(values = "red", name = '') + 
    geom_hline(yintercept = 0)

```{r mpa-fit-plot, fig.cap="Estimated MPA effects using PISCO data from inside MPAs only. The base model uses data from both inside and outside MPAs. These results show that the same general trend holds when only using data from inside the MPAs, though the model estimates greater probability of higher positive effects with only MPA data, compared to the model fit using all the data.",results = "hide"}

ggplot() +
 tidybayes::stat_halfeye(
    data = val_runs$did_results[[4]],
    aes(year, did, color = "Empirical Estimate"),
    .width = c(0.5, 0.95),
    slab_alpha = 0.95  ) +
  scale_y_continuous(name = "MPA Effect") +
  scale_x_discrete(name = element_blank()) +
  theme(legend.position = "top") +
  scale_fill_brewer(name = "Simulation Quantiles") + 
  scale_color_manual(values = "red", name = '') + 
  geom_hline(yintercept = 0)

```{r, fig.cap="Estimated MPA effects using PISCO data from outside MPAs only. The base model uses data from both inside and outside MPAs. These results show that the same general trend holds when only using data from outside the MPAs, though the model estimates less probability of high positive effects witout MPA data, compared to the model fit using all the data.",results = "hide"}

ggplot() +
 tidybayes::stat_halfeye(
    data = val_runs$did_results[[6]],
    aes(year, did, color = "Empirical Estimate"),
    .width = c(0.5, 0.95),
    slab_alpha = 0.95  ) +
  scale_y_continuous(name = "MPA Effect") +
  scale_x_discrete(name = element_blank()) +
  theme(legend.position = "top") +
  scale_fill_brewer(name = "Simulation Quantiles") + 
  scale_color_manual(values = "red", name = '') + 
    geom_hline(yintercept = 0)

```{r no-filter-plot, fig.cap="The base model only includes consistently observed species of finfish. For this run, we include species that have been seen in at least two years (down from fifteen). The results remain essentially unchanged.",results = "hide"}
no_filter <- model_runs %>%
  slice(1) %>%
  left_join(abundance_data, by = "data_source") %>%
  mutate(
    did_fit = pmap(
      list(
        data = data,
        data_to_use = data_to_use,
        data_source = data_source
      ),
      estimate_did,
      min_classcode_years = 2,
      iter = 2500,
      site_data = site_data,
      cdfw_catches = cdfw_catches,
      life_history_data = life_history_data
    )
  )

ggplot() +
 tidybayes::stat_halfeye(
    data = no_filter$did_fit[[1]]$did_results,
    aes(year, did, color = "Empirical Estimate"),
    .width = c(0.5, 0.95),
    slab_alpha = 0.95  ) +
  scale_y_continuous(name = "MPA Effect") +
  scale_x_discrete(name = element_blank()) +
  theme(legend.position = "top") +
  scale_fill_brewer(name = "Simulation Quantiles") + 
  scale_color_manual(values = "red", name = '') + 
    geom_hline(yintercept = 0)

```{r no-bin-plot, fig.cap="For ease of interpretation and model convergence, our default model estimates MPA effects in three-year bins. For this run we estimate the effect annually, relative to the year 2000. Results show the same overall trend and magnitude reported in our base model.", results = "hide", eval = FALSE}
no_bin <- model_runs %>%
  slice(1) %>%
  left_join(abundance_data, by = "data_source") %>%
  mutate(
    did_fit = pmap(
      list(
        data = data,
        data_to_use = data_to_use,
        data_source = data_source
      ),
      estimate_did,
      iter = 2500,
      first_year = 2001,
      bin_width = 1,
      site_data = site_data,
      cdfw_catches = cdfw_catches,
      life_history_data = life_history_data
    )
  )

ggplot() +
 tidybayes::stat_halfeye(
    data = no_bin$did_fit[[1]]$did_results,
    aes(year, did, color = "Empirical Estimate"),
    .width = c(0.5, 0.95),
    slab_alpha = 0.95  ) +
  scale_y_continuous(name = "MPA Effect") +
  scale_x_discrete(name = element_blank()) +
  theme(legend.position = "top") +
  scale_fill_brewer(name = "Simulation Quantiles") + 
  scale_color_manual(values = "red", name = '') + 
    scale_x_discrete(guide = guide_axis(n.dodge = 2))

```{r mean-fit, fig.cap="Our base run estimates the effect of MPAs on total biomass densities of targeted finfish. For this run, we instead estimate the effect on mean biomass densities of targeted species. Results show the same pattern as our base run. "}
mean_fit <- model_runs %>%
  slice(1) %>%
  left_join(abundance_data, by = "data_source") %>%
  mutate(
    did_fit = pmap(
      list(
        data = data,
        data_to_use = data_to_use,
        data_source = data_source
      ),
      estimate_did,
      fit_mean = TRUE,
      iter = 2500,
      site_data = site_data,
      cdfw_catches = cdfw_catches,
      life_history_data = life_history_data
    )
  )

ggplot() +
 tidybayes::stat_halfeye(
    data = mean_fit$did_fit[[1]]$did_results,
    aes(year, did, color = "Empirical Estimate"),
    .width = c(0.5, 0.95),
    slab_alpha = 0.95  ) +
  scale_y_continuous(name = "MPA Effect") +
  scale_x_discrete(name = element_blank()) +
  theme(legend.position = "top") +
  scale_fill_brewer(name = "Simulation Quantiles") + 
  scale_color_manual(values = "red", name = '') + 
    scale_x_discrete(guide = guide_axis(n.dodge = 2))

 exclude zeros

```{r no-zero-fit, fig.cap="Our base run estimates the effect of MPAs on total biomass densities of targeted finfish. For this run, we instead estimate the effect on mean biomass densities of targeted species. Results show the same pattern as our base run. "}

nozero_fit <- model_runs %>%
  slice(1) %>%
  left_join(abundance_data, by = "data_source")

nozero_fit$data[[1]] <- nozero_fit$data[[1]] %>% filter(any_seen == TRUE)

nozero_fit <-  nozero_fit%>%
  mutate(
    did_fit = pmap(
      list(
        data = data,
        data_to_use = data_to_use,
        data_source = data_source
      ),
      estimate_did,
      iter = 2500,
      site_data = site_data,
      cdfw_catches = cdfw_catches,
      life_history_data = life_history_data
    )
  )

ggplot() +
 tidybayes::stat_halfeye(
    data = nozero_fit$did_fit[[1]]$did_results,
    aes(year, did, color = "Empirical Estimate"),
    .width = c(0.5, 0.95),
    slab_alpha = 0.95  ) +
  scale_y_continuous(name = "MPA Effect") +
  scale_x_discrete(name = element_blank()) +
  theme(legend.position = "top") +
  scale_fill_brewer(name = "Simulation Quantiles") + 
  scale_color_manual(values = "red", name = '') + 
    scale_x_discrete(guide = guide_axis(n.dodge = 2))

```
 selection on observables

 ### Transect Level Difference-in-difference Model

Our base model aggregated the data to the level of total biomass densities of targeted and non-targeted species. This serves to average out some of the sampling noise from the indivudla transect level, and we would argue is the appropriate level of estimation for many key policy questions. However, this approach does oversimplify the data collection process, treating the total biomass densities at the year/site level as estimated without error.
